Reproducible Workflow on a Public Cloud for Computational Fluid Dynamics

### 2019-04-09 tex/MesnardoBarba2019.tex
Add small edits to workflow section

To overcome the so-called ``dependency hell'' and facilitate reproducibility and portability, we use the container technology provided by the open-source project Docker.
A container represents an isolated user space where application programs run directly on the operating system's kernel (with limited access to its resources).
In contrast with virtual machines, containers do not include a full operating system, making them lighter and faster.
Containers allow re-creating the same runtime environment of an application (including all its dependencies) from one machine to another. 
They empower researchers to share pre-built images of their software stack, ameliorating one of the biggest pain points for reproducible research: building the software in a different machine. 
The majority of enterprise developers today are familiar with Docker container technology (first released six years ago). 
But in academic settings, many are still unaware of its value. 
We present this section as an overview for research software developers unfamiliar with Docker, but comfortable with scripting, version control, and distributed collaboration.

A container is an instance of an image.
The developer builds this image on a local machine and pushes it to a public registry to share it with other users.
Users pull the image from the public registry,---in our case, DockerHub---and create containers out of it.
To create a Docker image, the developer writes a \texttt{Dockerfile}.: an ASCII file containing commands that tell Docker what to do.
For example, we create a new image from a base image using the keyword \texttt{FROM}.
We then write the different commands to build a multi-layered runtime environment of the application that includes all its dependencies.
Once we have built the image on the local machine, we push it to a repository on the DockerHub registry: a place to store and retrieve Docker images.
Listing \ref{lst:docker_build_push} provides the command lines to build (\texttt{docker build}) and push (\texttt{docker push}) an image of our CFD software (\texttt{barbagroup/petibm:0.4-GPU-IntelMPI-ubuntu}) that we used to obtain some of the results presented in the next section.
Here, \texttt{CLOUDREPRO} is an environment variable set to the local path of the GitHub repository for this paper, \texttt{cloud-repro}\footnote{\url{https://github.com/barbargroup/cloud-repro}} cloned on the user's machine.

@@

To run computational jobs on Microsoft Azure, we use several tools that facilitate creating and managing virtual machines with compute nodes, and submitting jobs to those nodes.
We use a service called Azure Batch that leverages Microsoft Azure at no extra cost, relieving the user from manually creating, configuring, and managing an HPC-capable cluster of cloud nodes, including virtual machines, virtual networks, job and task scheduling infrastructure.
Azure Batch works with both embarrassingly parallel workloads and tightly-coupled MPI jobs (the latter being the case of our CFD software).
To use Azure Batch, we first need to configure a workspace on Microsoft Azure.
This can be done either via the Azure Portal in a web browser or from a local terminal using the open-source tool Azure CLI.\footnote{Azure CLI (version 2.0.57): \url{https://github.com/Azure/azure-cli}}
We prefer to use the command-line solution (program \texttt{az}), as it allows us to keep track of the steps done to configure the Azure workspace (see Listing \ref{lst:az_configure}).
First, we set the subscription we want to use (let's call it \texttt{reprosubscription}).
Next, we create a resource group, \texttt{reprorg}, located in this case in the East US region, which will contain all the Azure resources.
We create an Azure Storage account, \texttt{reprostorage}, in the resource group,
and we also need to create an Azure Batch account, \texttt{reprobatch} and associate it to the storage account.
Finally, we create a fileshare (of size $100$ Gb) in the storage.

@@

Once the configuration files are written, we invoke Batch Shipyard through the \texttt{shipyard} program on our local machine.
The folder \texttt{examples/snake2d2k35/config\_shipyard} in the repository accompanying this paper contains an example YAML files to create a pool of two \texttt{NC24r} compute nodes (featuring K80 GPUs and InfiniBand network).
Listing \ref{lst:shipyard_run} shows the commands to run from your local terminal to create a pool of compute nodes on Azure Batch and submit jobs to it.
During the pool creation (\texttt{shipyard pool add}), the Docker image of our CFD application is pulled from the registry to the virtual machines.
With \texttt{shipyard data ingress}, we can upload input files to the compute nodes.
Once created, we submit jobs to the pool (\texttt{shipyard jobs add}); the tasks for a job will start automatically.

@@

Once the simulations are done (i.e., the job tasks are complete), we delete the jobs and the pool (Listing \ref{lst:shipyard_del}).
The output of the computation is now stored in the fileshare in our Azure Storage account.
We can download the data to our local machine to perform additional post-processing steps.

>>>>

To overcome the so-called ``dependency hell'' and facilitate reproducibility and portability, we use the container technology provided by the open-source project Docker.
A container represents an isolated user space where application programs run directly on the operating system's kernel of the host (with limited access to its resources).
In contrast with virtual machines, containers do not include a full operating system, making them lighter and faster.
Containers allow re-creating the same runtime environment of an application (including all its dependencies) from one machine to another. 
They empower researchers to share pre-built images of their software stack, ameliorating one of the biggest pain points for reproducible research: building the software in a different machine. 
The majority of enterprise developers today are familiar with Docker container technology (first released six years ago). 
But in academic settings, many are still unaware of its value. 
We present this section as an overview for research software developers unfamiliar with Docker, but comfortable with scripting, version control, and distributed collaboration.

A container is an instance of an image.
The developer builds this image on a local machine and pushes it to a public registry to share it with other users.
Users pull the image from the public registry,---in our case, DockerHub---and create containers out of it.
To create a Docker image, the developer writes a \texttt{Dockerfile}: an ASCII file containing instructions that tell Docker what to do.
For example, we start building a new image from a base image using the keyword \texttt{FROM}.
We then write the different shell instructions (with the \texttt{RUN} instruction) to build a multi-layered runtime environment of the application that includes all its dependencies.
Once we have built and tested the image on the local machine, we push it to a repository on the DockerHub registry: a place to store and retrieve Docker images.
Listing \ref{lst:docker_build_push} provides the command lines to build (\texttt{docker build}) and push (\texttt{docker push}) an image of our CFD software (\texttt{barbagroup/petibm:0.4-GPU-IntelMPI-ubuntu}) that we used to obtain some of the results presented in the next section.
Here, \texttt{CLOUDREPRO} is an environment variable set to the local path of the GitHub repository for this paper, \texttt{cloud-repro}\footnote{\url{https://github.com/barbargroup/cloud-repro}} cloned on the user's machine.

@@

To run computational jobs on Microsoft Azure, we use several tools that facilitate creating and managing virtual machines with compute nodes, and submitting jobs to those nodes.
We use a service called Azure Batch that leverages Microsoft Azure at no extra cost, relieving the user from manually creating, configuring, and managing an HPC-capable cluster of cloud nodes, including virtual machines, virtual networks, job and task scheduling infrastructure.
Azure Batch works with both embarrassingly parallel workloads and tightly-coupled MPI jobs (the latter being the case of our CFD software).
To use Azure Batch, we first need to configure a workspace on Microsoft Azure.
This can be done either via the Azure Portal in a web browser or from a local terminal using the open-source tool Azure CLI.\footnote{Azure CLI (version 2.0.57): \url{https://github.com/Azure/azure-cli}}
We prefer to use the command-line solution (program \texttt{az}), as it allows us to keep track of the steps done to configure the Azure workspace (see Listing \ref{lst:az_configure}).
First, we set the Azure subscription we want to use (let's call it \texttt{reprosubscription}).
Next, we create a resource group, \texttt{reprorg}, located in this case in the East US region, which will contain all the Azure resources.
We create an Azure Storage account, \texttt{reprostorage}, in the resource group, as well as an Azure Batch account, \texttt{reprobatch}, associated to the storage account.
Finally, we create a fileshare (in this case of size $100$ GB) in the storage.

@@

Once the configuration files are written, we invoke Batch Shipyard through the \texttt{shipyard} program on our local machine.
The folder \texttt{examples/snake2d2k35/config\_shipyard} in the repository accompanying this paper contains an example YAML files to create a pool of two \texttt{NC24r} compute nodes (featuring K80 GPUs and using InfiniBand network).
Listing \ref{lst:shipyard_run} shows the commands to run from your local terminal to create a pool of compute nodes on Azure Batch and submit jobs to it.
During the pool creation (\texttt{shipyard pool add}), the Docker image of our CFD application is pulled from the registry to the virtual machines.
With \texttt{shipyard data ingress}, we can upload input files to the compute nodes.
Once created, we submit jobs to the pool (\texttt{shipyard jobs add}); the tasks for a job will start automatically.

@@

Once the simulations are done (i.e., the job tasks are complete), we delete the jobs and the pool (Listing \ref{lst:shipyard_del}).
The output of the computation is now stored in the fileshare in our Azure Storage account.
We can download the data to our local machine to perform additional post-processing steps (such as flow visualizations).
